Concurrent Programming with Harmony

### 2020-08-10 book/paper.tex
Updated example with graphics

\begin{code}
\begin{verbatim}
$ harmony triangle.hny
#states = 13 diameter = 1
no issues found
$
\end{verbatim}
\end{code}

@@

In Harmony, constants have a default value,
but those can be overridden on the command
line using the \texttt{-c} option.
\index{constant}%
For example, if you want to test the code for \texttt{N = 100}, run:
\begin{code}
\begin{verbatim}
$ harmony -c N=100 triangle.hny
#states = 103 diameter = 1
no issues found
$
\end{verbatim}
\end{code}

@@

\begin{figure}
\begin{code}
\begin{verbatim}
Up.hny:1 def incrementer(self):
    0 Jump 34
    1 Frame incrementer(self)
Up.hny:2     count = count + 1;
    2 Load count
    3 Push 1
    4 2-ary +
    5 Store count
Up.hny:3     done[self] = True;
    6 PushAddress done
    7 LoadVar self
    8 Address 2
    9 Push True
   10 Store
   11 Return
\end{verbatim}
\end{code}
\caption{The first part of the HVM bytecode corresponding to \autoref{fig:inc}.}
\label{fig:inccode}
\end{figure}

@@

\begin{figure}
\begin{code}
\begin{verbatim}
#states = 79 diameter = 5
==== Safety violation ====
__init__/()   [0,34-49]     50 { count=0, done=[False, False] }
incrementer/0 [1-4]          5 { count=0, done=[False, False] }
incrementer/1 [1-10]        11 { count=1, done=[False, True]  }
incrementer/0 [5-10]        11 { count=1, done=[True,  True]  }
main/() [13-17,20-24,26-31] 31 { count=1, done=[True,  True]  }
>>> Harmony Assertion failed : 1
Open file://.../harmony.html for more information
\end{verbatim}
\end{code}
\caption{The text output of running Harmony on \autoref{fig:inc}.}
\label{fig:incoutput}
\end{figure}

@@

Compiling the code in \autoref{fig:inc} results in the HVM bytecode
listed in \autoref{fig:inccode}.
You can obtain this code by invoking \texttt{harmony} with the \texttt{-a} flag
like so:
\begin{code}
\begin{verbatim}
harmony -a Up.hny
\end{verbatim}
\end{code}
Each process in the HVM is predominantly a \emph{stack machine},
\index{stack machine}%
but it also has a register.
All instructions are atomically executed.
Most instructions pop values from the stack or push values onto the stack.
At first there is one process, with name tag \texttt{\_\_init\_\_/()},
that initializes the state.
It starts executing at instruction 0 and keeps executing until the last
execution in the program.
In this case, the first instruction is a \texttt{JUMP} instruction that sets the
program counter to 34 (skipping over the code for the methods).
At program counter 1 is the code for the \texttt{incrementer} method.
All methods start with a \texttt{Frame} instruction and end with a \texttt{Return}
instruction.

@@

In this case, Harmony reports a safety violation, in particular it
reports that the assertion on line 9 failed and that
\textit{count} has value 1 instead of 2.
Next it reports an execution that failed this assertion.
The program got to the faulty execution in 5 ``steps.''
The output has
four columns:
\begin{enumerate}
\item The name tag of the process;
\item The sequence of program counters of the HVM instructions that the process executed;
\item The current program counter of the process;
\item The contents of the shared memory.
\end{enumerate}

The first step is initialization.
It sets shared variable \textit{count} to 0 and
shared variable \textit{done} to \texttt{[False, False]}.
If we look at the next three steps, we see that:
\begin{itemize}
\item Process \texttt{implementer/0}
executed instructions 1 through 4, loading the value of
\textit{count} but stopping just before storing 1 into \textit{count};
\item Process \texttt{implementer/1}
executed instructions 1 through 10, storing 1 into
into \textit{count} and storing \texttt{True} into \textit{done}[1];
\item Process \texttt{implementer/0}
continues execution, storing value 1 into \textit{count}
and storing \texttt{True} into \textit{done}[0].
\end{itemize}
Finally, process \texttt{main()} runs and finds the problem.
This makes precise the concurrency issue that we encountered.

>>>>

\begin{code}
\begin{verbatim}
$ harmony triangle.hny
#states = 13 diameter = 1
#components: 13
no issues found
Open file:///.../harmony.html for more information
$
\end{verbatim}
\end{code}

@@

In Harmony, constants have a default value,
but those can be overridden on the command
line using the \texttt{-c} option.
\index{constant}%
For example, if you want to test the code for \texttt{N = 100}, run:
\begin{code}
\begin{verbatim}
$ harmony -c N=100 triangle.hny
#states = 103 diameter = 1
#components: 103
no issues found
Open file:///.../harmony.html for more information
$
\end{verbatim}
\end{code}

@@

\begin{figure}
\begin{code}
\begin{verbatim}
code/Up.hny:1 def incrementer(self):
   0 Jump 25
   1 Frame incrementer self
code/Up.hny:2     count = count + 1;
   2 Load count
   3 Push 1
   4 2-ary +
   5 Store count
code/Up.hny:3     done[self] = True;
   6 PushAddress done
   7 LoadVar self
   8 Address 2
   9 Push True
   10 Store
   11 Return
code/Up.hny:5 def main():
   12 Jump 25
   13 Frame main ()
code/Up.hny:6     await all(done);
   14 Load done
   15 1-ary all
   16 JumpCond False 14
code/Up.hny:7     assert count == 2, count;
   17 AtomicInc
   18 Load count
   19 Push 2
   20 2-ary ==
   21 Load count
   22 Assert2
   23 AtomicDec
   24 Return
code/Up.hny:9 count = 0;
   25 Push 0
   26 Store count
code/Up.hny:10 done = [ False, False ];
   27 Push [False, False]
   28 Store done
code/Up.hny:11 spawn incrementer(0);
   29 Push 0
   30 Dup
   31 Push PC(1)
   32 Spawn
\end{verbatim}
\end{code}
\caption{The first part of the HVM bytecode corresponding to \autoref{fig:inc}.}
\label{fig:inccode}
\end{figure}

@@

\begin{figure}
\begin{code}
\begin{verbatim}
#states = 71 diameter = 5
==== Safety violation ====
__init__/()   [0,25-41] 41 { count: 0, done: [False, False] }
incrementer/0 [1-4]      5 { count: 0, done: [False, False] }
incrementer/1 [1-11]    11 { count: 1, done: [False, True ] }
incrementer/0 [5-11]    11 { count: 1, done: [True,  True ] }
main/()       [13-22]   22 { count: 1, done: [True,  True ] }
>>> Harmony Assertion failed : 1
Open file://.../harmony.html for more information
\end{verbatim}
\end{code}
\caption{The text output of running Harmony on \autoref{fig:inc}.}
\label{fig:incoutput}
\end{figure}

@@

Compiling the code in \autoref{fig:inc} results in the HVM bytecode
listed in \autoref{fig:inccode}.
You can obtain this code by invoking \texttt{harmony} with the \texttt{-a} flag
like so:
\begin{code}
\begin{verbatim}
harmony -a Up.hny
\end{verbatim}
\end{code}
Each process in the HVM is predominantly a \emph{stack machine},
\index{stack machine}%
but it also has a register.
All instructions are atomically executed.
Most instructions pop values from the stack or push values onto the stack.
At first there is one process, with name tag \texttt{\_\_init\_\_/()},
that initializes the state.
It starts executing at instruction 0 and keeps executing until the last
execution in the program.
In this case, the first instruction is a \texttt{JUMP} instruction that sets the
program counter to 25 (skipping over the code for the methods).
At program counter 1 is the code for the \texttt{incrementer} method.
All methods start with a \texttt{Frame} instruction and end with a \texttt{Return}
instruction.

@@

In this case, Harmony reports a safety violation, in particular it
reports that the assertion and that
\textit{count} has value 1 instead of 2.
Next it reports an execution that failed this assertion.
The program got to the faulty execution in 5 ``steps.''
The output has
four columns:
\begin{enumerate}
\item The name tag of the process;
\item The sequence of program counters of the HVM instructions that the process executed;
\item The current program counter of the process;
\item The contents of the shared memory.
\end{enumerate}

The first step is initialization.
It sets shared variable \textit{count} to 0 and
shared variable \textit{done} to \texttt{[False, False]}.
If we look at the next three steps, we see that:
\begin{itemize}
\item Process \texttt{implementer/0}
executed instructions 1 through 4, loading the value of
\textit{count} but stopping just before storing 1 into \textit{count};
\item Process \texttt{implementer/1}
executed instructions 1 through 11, storing 1 into
into \textit{count} and storing \texttt{True} into \textit{done}[1];
\item Process \texttt{implementer/0}
continues execution, executing instructions 5 through 11
storing value 1 into \textit{count}
and storing \texttt{True} into \textit{done}[0].
\end{itemize}
Finally, process \texttt{main()} runs and finds the problem.
This makes precise the concurrency issue that we encountered.
